Generating Empirically Optimized Numerical Software from MATLAB Prototypes

### 2008-09-11 doc/papers/cgo08/paper.tex
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@300 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb
current gdoc version
git-svn-id: https://svn.mcs.anl.gov/repos/performance/orio@347 e7ad4b5f-b827-0410-872f-f7f4bc3d1efb

% CGO 09: http://www.cgo.org/
% svn co https://svn.mcs.anl.gov/repos/performance/orio/doc/papers/cgo08
% Max 22 pages, double-spaced, 11pt or larger font
% Abstract: Sep. 12, 2008
% Full paper: Sep. 19, 2008

\documentclass[11pt]{article}

\usepackage{url}
\usepackage{html}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{times}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}

@@

\author{
Elizabeth Jessup,$^1$ Boyana Norris,$^2$ Albert Hartono,$^3$ and Jeremy Siek$^4$\\
\\
\vspace{-.05in}
{\small $^1$University of Colorado at Boulder Dept. of Computer Science, Boulder, CO~80302}\\
\vspace{-.05in}
{\small \url{jessup@cs.colorado.edu}}\\
\vspace{-.05in}
{\small $^2$Argonne National Laboratory Mathematics and Computer Science Division, Argonne, IL~60439}\\
\vspace{-.05in}
{\small \url{norris@mcs.anl.gov}}\\
\vspace{-.05in}
{\small $^3$Ohio State University Dept. of Computer Science and Engineering, Columbus, OH 43210}\\
\vspace{-.05in}
{\small \url{hartonoa@cse.ohio-state.edu}}\\
\vspace{-.05in}
{\small $^4$University of Colorado at Boulder Dept. of Electrical and Computer Engineering, Boulder, CO~80302}\\
\vspace{-.05in}
{\small \url{jeremy.siek@colorado.edu}}
}

@@

\abstract{TODO}
\doublespacing
\section{Introduction}
\label{sec:intro}

@@

\subsection{A {MATLAB} Compiler}
\label{sec:matlab}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\textwidth]{figures/refine-mv-dot.png}
\caption{The refinement of a matrix-vector product to a set of inner products and the refinement of inner products to scalar multiplications and additions.}
\label{fig:refine-mv-dot}
\end{figure}

@@

@@

\subsection{WAXPBY $v = Ax + By$}
perhaps add another term

@@

%for (i=0; i<=A_nrows-1; i++)
%  for (j=0; j<=A_ncols-1; j++)
%    B[i][j] = A[i][j] + u1[i]*v1[j] + u2[i]*v2[j];
%
%for (i=0; i<=A_ncols-1; i++)
%  for (j=0; j<=A_nrows-1; j++)
%    x[i] = x[i] + b * B[j][i]*y[j];
%
%
%for (i=0; i<=A_ncols-1; i++)
%  x[i] = x[i] + z[i];
%
%for (i=0; i<=A_nrows-1; i++)
%  for (j=0; j<=A_ncols-1; j++)
%    w[i] = w[i] + a * B[i][j]*x[j];

@@

% Placeholder for the new results, the loops tuned for the figure above were almost identical to the ones generated by the MATLAB compiler

\subsection{GESUMMV $y = a * (A * x) + b * (B * x)$}

\subsection{Bicgkernel}

@@

extend beyond composed BLAS
\singlespacing
\section*{References}
\bibliographystyle{siam}
\bibliography{paper}
\end{document}

>>>>

% CGO 09: http://www.cgo.org/
% svn co https://svn.mcs.anl.gov/repos/performance/orio/doc/papers/cgo08
% Max 22 pages, double-spaced, 11pt or larger font
% Abstract: Sep. 12, 2008
% Full paper: Sep. 19, 2008
\documentclass[11pt]{article}

\usepackage{url}
\usepackage{html}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}

@@

\author{
Elizabeth Jessup,$^1$ Boyana Norris,$^2$ Albert Hartono,$^3$ and Jeremy Siek$^4$\\
\\
\vspace{-.05in}
{\small $^1$University of Colorado at Boulder Dept. of Computer Science, Boulder, CO~80309}\\
\vspace{-.05in}
{\small \url{jessup@cs.colorado.edu}}\\
\vspace{-.05in}
{\small $^2$Argonne National Laboratory Mathematics and Computer Science Division, Argonne, IL~60439}\\
\vspace{-.05in}
{\small \url{norris@mcs.anl.gov}}\\
\vspace{-.05in}
{\small $^3$Ohio State University Dept. of Computer Science and Engineering, Columbus, OH~43210}\\
\vspace{-.05in}
{\small \url{hartonoa@cse.ohio-state.edu}}\\
\vspace{-.05in}
{\small $^4$University of Colorado at Boulder Dept. of Electrical and Computer Engineering, Boulder, CO~80309}\\
\vspace{-.05in}
{\small \url{jeremy.siek@colorado.edu}}
}

@@

\abstract{TODO}

\doublespacing
\section{Introduction}
\label{sec:intro}

@@

\subsection{A MATLAB Compiler}
\label{sec:matlab}

@@

\begin{figure}[hbtp]
  \centering
  \includegraphics[width=\textwidth]{figures/refine-mv-dot.png}

  \caption{The refinement of a matrix-vector product to a set of inner products and the refinement of inner products to scalar multiplications and additions.}
  \label{fig:refine-mv-dot}
\end{figure}

@@

In this section we present performance results for a number of composed BLAS operations. We performed experiments on an dual quad-core 2/88 Xeon with 2.88 GHz, 16GB DDR2 FBDIMM

@@

$x = w + y + z$

\subsection{ATAX}

$y = A' * (A * x)$

\begin{figure}[htp]
\centering
\includegraphics[width=.8\textwidth]{figures/atax.png}
\caption{Wall-clock time (in seconds) for the ATAX operation using different code versions.}
\label{fig:atax}
\end{figure}

@@

\begin{figure}[htb]
\centering
\scriptsize
\begin{verbatim}
for (i=0; i<=A_nrows-1; i++)
  for (j=0; j<=A_ncols-1; j++)
   B[i][j] = A[i][j] + u1[i]*v1[j] + u2[i]*v2[j];

for (i=0; i<=A_ncols-1; i++)
  for (j=0; j<=A_nrows-1; j++)
    x[i] = x[i] + b * B[j][i]*y[j];

for (i=0; i<=A_ncols-1; i++)
  x[i] = x[i] + z[i];

for (i=0; i<=A_nrows-1; i++)
  for (j=0; j<=A_ncols-1; j++)
    w[i] = w[i] + a * B[i][j]*x[j];
\end{verbatim}
\caption{GEMVER C code.}
\label{fig:gemver}
\end{figure}

%--- Sequential: seconds (dynamic arrays) ---
%cmatlab= [0.038, 0.157, 0.352, 0.626, 0.976, 1.406, 1.909, 2.494, 3.143, 3.886]
%orio= [0.029, 0.118, 0.263, 0.480, 0.759, 1.075, 1.447, 1.899, 2.338, 2.978]
%blas= [0.064, 0.265, 0.598, 1.091, 1.711, 2.456, 3.340, 4.345, 5.480, 6.840]
%mkl=  [0.073, 0.147, 0.331, 0.591, 0.923, 1.326, 1.803, 2.350, 2.975, 3.676]
%atlas=[0.039, 0.163, 0.401, 0.796, 1.225, 1.916, 2.582, 3.381, 4.210, 4.756]
%
%--- Parallel: seconds (dynamic arrays) --- N=10,000, cores=1-8
%cmatlab= [0.976, 0.976, 0.976, 0.976, 0.976, 0.976, 0.975, 0.976]
%orio= [0.754, 0.738, 0.666, 0.649, 0.664, 0.664, 0.647, 0.668]

@@

\begin{figure}[htp]
\centering
\includegraphics[width=.8\textwidth]{figures/gemver.png}
\caption{Wall-clock time (in seconds) for the GEMVER operation using different code versions.}
\label{fig:gemver}
\end{figure}

\subsection{GESUMMV}
$y = a * (A * x) + b * (B * x)$

\begin{figure}[htp]
\centering
\includegraphics[width=.8\textwidth]{figures/gesummv.png}

\caption{Wall-clock time (in seconds) for the GESUMMV operation using different code versions.}
\label{fig:gesummv}
\end{figure}
\subsection{BiCG Kernel}

$q = A * p$, $s = A' * r$

\begin{figure}[htp]
\centering

\includegraphics[width=.8\textwidth]{figures/bicgkernel.png}
\caption{Wall-clock time (in seconds) for the BiCG Kernel operation using different code versions.}
\label{fig:bicgkernel}
\end{figure}

%NOTES about the experiments

%Results already obtained:
% -gemver (orio/testsuite/gemver/cookie directory)
% -gesummv (orio/testsuite/gesummv/cookie directory)
% -bicgkernel (orio/testsuite/bicgkernel/cookie directory)
% -atax (orio/testsuite/atax/cookie directory)

%* All results were obtained on the cookie machine (eight-core Intel Xeon)

%* Compiler used: icc 10.1
%* Optimization flags used: -O3 (-parallel for parallel matlab code) (-openmp for parallel orio code)

%* We compare two types of codes:
%  1. matlab
%  2. orio
%1. 'matlab' code is the original code produced by the matlab c compiler.
%2. 'orio' is obtained first from the Pluto-generated code (use loop fusion, NO loop tiling,
%   NO unrolling, NO parallelization, NO vectorization), then was decorated with
%   performance annotations before being tuned by Orio.

%* Each code has two versions: one with static arrays, and another with dynamic arrays.

%* Conversion from/to static arrays to/from dynamic arrays was done manually.

%* I run Pluto first to maximally fuse all loops (using polycc <file>
%--maxfuse option) and see the resulting fused loop(s). And then tune the
%fused loop using Orio. Sometimes, the fused loop gets very complicated
%because of Pluto's various transformations (e.g. skewing). In that case, I
%don't use Pluto's loop fusion. For example, gemver and gesummv results use
%Pluto's loop fusion, but bicgkernel doesn't use Pluto's loop fusion at
%all.

%* All experiment files are in <kernel-name>/cookie directory

%* All generated codes (the base Matlab codes, and the Orio-tuned codes),
%results, script for running the experiments are in
%<kernel-name>/cookie/exp

%* I tuned 4 different versions for each kernel:
% 1. sequential (static arrays)
% 2. sequential (dynamic arrays)
% 3. parallel (static arrays)
% 4. parallel (dynamic arrays)

%The annotated source code for 1 and 3 is in <kernel-name>.src1.c
%The annotated source code for 3 and 4 is in <kernel-name>.src2.c

%The final code for 1 and 2 is in <kernel-name>.orio.seq.c
%The final code for 3 and 4 is in <kernel-name>.orio.par.c

%* <kernel-name>.matlab.c contains the original Matlab code (for both
%static and dynamics arrays)

%* The best performance parameters for each version can be seen in
%results.txt file.

@@

extend beyond composed BLAS

Ultimately, the tuning process should result in more than one version of the computation since different versions are optimal for different problem sizes.

\singlespacing
\bibliographystyle{siam}
\bibliography{paper}
\end{document}
